Comprehensive Algorithmic Resilience for Scientific Applications

### 2013-01-23 paper.tex
Changes in text; change the facet names on figs 9~14

\subsection{Alternating Direction Method of Multipliers (ADDR)}
\label{sec:apps:lasso}
The ADDR method solves under-unconstrained linear problems $Ax=b$ for $x$ ($A$ has fewer rows than columns) while minimizing the function $f(x)$.
This is an iterative algorithm uses 64-bit precision and that spends most of its time in the following linear algebra operations, matrix-matrix multiplication, matrix-vector multiplication, rank-k update and Cholesky factorization, using the implementations of these routines from the GNU Scientific Library~\cite{gsl:2011}.
We focus on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
In our experiments we consider linear systems of size $\{40, 80, 120, ..., 4000\} \times 500$ that are generated by sampling a normal distribution.
According to our experiments, the running time of this solver is not dependent on the kind of distribution the inputs are sampled from.

@@

The DRC program generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
The inputs are stored in Pulse Code Modulation (PCM) format, which is an array of 32-bit floating point numbers representing the samples at each sampling time step and the computations are performed using 32-bit precision.
Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform and an DRC-internal implementation of Finite Impulse Response filter generation.
This program accepts user-controllable configurations to decide the width and number of iterations of FFT transforms it uses to perform various passes on the input file, such as windowing, deconvolution, pre- and post-filtering.
The input used in this paper is an audio file of size 768 kilobytes.

\subsection{Hattrick}
\label{sec:apps:hattrick}
The Hattrick application simulates the motion of bodies under the effects of gravity using 64-bit precision.
It is designed to help discover extra-solar planets by inferring their existence from Transit Timing Variations, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solver the system's equations of motion.
Input files are four different planetary systems. The first, third and fourth input files simulate one star and one planet and the second simulates one star and two planets. The third and fourth inputs are the same planetary system with different lengths of simulation time.

@@

In this paper we consider a comprehensive resilience strategy that combines error detection and recovery.
Error detection techniques are discussed in Section~\ref{sec:res_tech:err_det}, with Section~\ref{sec:res_tech:err_det:algo} focusing on algorithmic techniques to detect errors in the numeric portions of application state, while Section~\ref{sec:res_tech:err_det:mem} focuses on protection of data structures.
Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart, which offers lower-overhead but higher-latency recovery, as well as replication of key variables, which reduces performance but enables fast recovery.
Finally, Section~\ref{sec:res_tech:eval} presents an experimental evaluation of the performance and accuracy properties of the resilient versions of the major routines used by our target applications, under a wide range of inputs and error injection rates.

@@

\subsubsection{Algorithmic Detection}
\label{sec:res_tech:err_det:algo}

@@

Each algorithm used by the above application was enhanced with an algorithm-specific checker that validated its results by exploiting some algorithmic identity.
Each checker computes some aggregate property of the algorithm's results using two different algorithm to produce vectors $v$ and $v'$ and their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$.
Each error detection causes the routine's execution to be restarted using the rollback mechanism described in Section~\ref{sec:res_tech:cr}.

\paragraph{Matrix-matrix multiplication (MMM)}
The MMM operation $A \cdot B$ is checked using a matrix vector multiplication (MVM), making use of this identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector.
Both sides compute the sums of the columns of matrix $A \cdot B$ using a single matrix-vector multiplication on the left and two such operations on the right.
Most errors in the computation of $A \cdot B$ will affect the sums of its columns and will thus be detected.
This check is efficient because the checker is asymptotically faster than the MMM algorithm, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
Since this is dense martix matrix multiplication, we set the check vector $x$ to contain all 1's.

\paragraph{Matrix-vector multiplication (MVM)}
The MVM algorithm $Ax=b$ is checked using a similar identity $(x^TA)x = x^T(Ax) = x^Tb$.
The product $x^TA$ is a checksum of the columns of matrix $A$ and when $x$ is the vector of 1's, the value used in our experiments, it is the sum of $A$'s columns.
Although the computation of $x^TA$ involves additions rather than multiplications, the complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM operation.
%While this cost can be reduced significantly by computing $x^TA$ once and reusing it in subsequent multiplications by the same matrix, we did not consider it in this paper because MVM took up little application execution time.
However, for applications such as ADDR where the MVM is applied to same matrix many times with different vectors, the column sums can be reused, amortizing the cost of this computation.
Our experimental evaluation below thus focuses on the overhead of the $x^T(Ax)$ computation, which has $O(n)$ complexity.

\paragraph{Symmetric Rank-K update (SYRK)}
The SYRK update is a special case of matrix-matrix multiplication $B = \alpha A \cdot A^T + \beta B$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is incremented into the $nxn$ matrix $B$.
Its results are checked using the same algorithm as MMM by checking the identity $B x = \alpha A \cdot (A^Tx) + \beta B x$, where $x$ is the vector of all 1's.
This check runs in time $O(n^2)$, as compared to the $O(n^3)$ cost of SYRK.

\paragraph{Cholesky Decomposition (CD)}
In this decomposition matrix $A$ is decomposed into $L \cdot L^T$ where $L$ is a lower-triangular matrix with a positive diagonal.
This routine is checked just like MMM by checking the identity $Ac = L \cdot (L^T x)$, which is cheaper than the $O(n^3)$ complexity of the deterministic CD algorithm.
The GSL implements an iterative algorithm that runs faster than $O(n^3)$ time but as shown in the experiments below, it is significantly slower than our checker.

\paragraph{Fast Fourier Transform (FFT)}
The FFT algorithm decomposes a function into a sum of sine waves of different frequencies: $f(x) = \sum_{n=0}^{N-1} x_n e^{-i2\pi k n / N}$ for some constant $k$.
We check the results of FFT by using Parseval's theorem, which states that $\sum_{n=0}^{N-1} \left| x[n] \right|^2 = \frac{1}{N} \sum_{k=0}^{N-1} \left| X[k] \right|^2$, where $x$ is the original function and $X$ is its transform. GSL implements optimized radix 2, 3, 4, 5, 6 and 7 subtransform routines as well as a general radix-N FFT routine. The subtransforms are optimized and consume $O(n log(n))$ time and the general one consumes $O(n^2)$ time. Therefore, GSL has a scheduling algorithm that factorizes the length of the input FFT, running the subtransforms whenever possible, using the fallback general radix-N routine only for the remaining factors.
The theorem applies to FFT of any radix.
This means that the total energy of the original function is preserved by the transform and provides an $O(n)$ time check for the $O(n log(n))$ FFT algorithm.

\paragraph{Finite Impulse Response Filter Generation (FIR)}
The FIR filter generation algorithm used in DRC generates a series of samples over the $sinc$ function $sinc(x)=\frac{sin(x)}{x} (sinc(x)=1\ when\ x\ is\ 0)$ and modulates it with a Blackman window.
This function has the property that $\int_{-\infty}^{\infty} sinc(x)dx = 1$. 
The Blackman window - modulated output of this routine preserves this same property for most of the time, making it possible to check the output by computing the sum and comparing it to 1.
Computing the sum requires $O(n)$ additions, as compared to the $O(n)$ trigonometric function evaluations of the non-iterative FIR generation algorithm.
%As shown below, the overhead of this checker is very small (less than 5\%).
%Since this algorithm has only output and no input and therefore it does not need to have to be reinforced with input backup as other routines do.

\paragraph{Runge-Kutta PDE Solver (RK)}
This algorithm uses the 4-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
It works by advancing the variable $x$ by steps of size $h$ and using estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
The RK method in GSL uses adaptive step-size control where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
If the difference between the results obtained with the two step sizes are larger than a high water-mark threshold, the algorithm switches to the smaller step size to maintain accuracy and resumes.
If it is smaller than a low water-mark threshold, the algorithm switches to a larger step size.
Because soft errors will usually cause significant differences between the two computations, the adaptive step size control algorithm is naturally fault tolerant.
It will react to such errors by rolling back, temporarily increasing its step size and then reverting to the original step size when it observes that this is safe.
As such, we utilize this algorithm to protect the computations within individual RK4 steps from soft errors.
However, computations in user-provided function derivative functions are not checked since their invariants are unknown.

@@

Since modern systems use 64-bit addressing the range of values that can be assumed by a given pointer is much larger than the amount of memory that can ever be assigned to an application.
This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
Once this address is accessed this violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or Bus Error signal.
This mechanism is cheap but suffers from a long delay between an error and its detection.

An alternative mechanism is to replicate key application variable and to compare the values of the replicas on each access to the variable.
While more expensive, this approach is effective if the number of key variables is small and the cost of rolling back computation is high.

@@

\subsection{Recovery via Checkpoint-Restart and Replication}
\label{sec:res_tech:cr}

When errors are detected we employ a recovery method based on checkpoint-restart, where routine state is recorded at key locations and whenever an algorithmic or memory error is detected we roll back to that point in its execution.
Rollbacks are implemented via the \texttt{sigsetjmp}/\texttt{siglongjmp} \cite{sigsetjmp:1997} routines.
\texttt{sigsetjmp} record the application's execution state (registers and stack pointer) at the time of the checkpoint.
On error detection \texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
This method is useful for rolling back to checkpoints recorded in the same function or in a caller function.
When execution returns to the \texttt{sigsetjmp} call the checkpointed portions of the routine's state are recovered and execution repeats.
% The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to negligible performance degradation.

Checkpoint data is protected using two different mechanisms:
Pointers and scalars are replicated such that the checkpoint contains two copies of each.
Arrays of floating point numbers are protected via a block-checksum-based error correcting approach:
By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, it detects the existence of errors and fixes them. 
This algorithm takes an array of numbers as input, divides it into blocks of $N \cdot N$ numbers and for each block it computes and stores the row sums and column sums, which are used for recovery.

The structure of each routine determines how checkpointing is deployed.
All routines except RK are monolithic code blocks that take a relatively large amount of data as input, operate on it and then return.
In contrast, RK takes in very little input and operates as a long sequence of steps, each of which involves a call to a user-provided derivative routine.
As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers.
The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed..
Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a fixed number of steps between checkpoints.
As our experiments in \ref{sec:res_tech:eval:rk} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.

@@

Having evaluated the resilience properties of the individual application routines and resilience techniques we now consider our three target applications ADDR, Hattrick and DRC.
Section~\ref{sec:eval:perf} reports on the impact of errors and resilience techniques on application performance. Section~\ref{sec:eval:acc} describes the distribution of errors in the results of these applications.
These results are based on a fault injection campaign that includes a range of fault injection rates, input sizes and configurations of resilience mechanisms, with 100 application runs for each combination of parameters.
This corresponds to a total of 66100 for lasso, 43200 for DRC, 102000 for Hattrick error injection experiments.

@@

\begin{figure*}[ht!]
\centering
\includegraphics[height=2.75in]{figs/Hattrick_EstdCost.png}
\vspace{-10pt}
\caption{Execution time of Hattrick relative to non-fault tolerant fault-free execution time. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
\vspace{-10pt}
\label{fig:Hattrick_EstdCost}
\end{figure*}

@@

\begin{figure*}[ht!]
\centering
%\includegraphics[height=3in]{figs/Hattrick_ImperfectRate.png}
\includegraphics[height=2.75in]{figs/Hattrick_ImperfectRate_log.png}
\caption{Probability of not satisfying the user-provided error bound for Hattrick. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
\label{fig:Hattrick_ImperfectRate}
\end{figure*}

>>>>

\subsection{Alternating Direction Method of Multipliers (ADDR)}
\label{sec:apps:lasso}
The ADDR method solves under-unconstrained linear problems $Ax=b$ for $x$ ($A$ has fewer rows than columns) while minimizing the function $f(x)$.
This is an iterative algorithm uses 64-bit precision and that spends most of its time in the following linear algebra operations, matrix-matrix multiplication, matrix-vector multiplication, rank-k update and Cholesky factorization, using the implementations of these routines from the GNU Scientific Library~\cite{gsl:2011}.
We focus on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
It is parallelized using MPI and our experiments focus on 4-processors runs of ADDR which computes linear systems of size $\{40, 80, 120, ..., 4000\} \times 500$ (40 to 4000 measurements with 500 variables each) and in the experiments below are denoted $S40$, $S80$, ..., $S4000$.
The values in $A$ and $b$ are generated by sampling a normal distribution with a mean of 0.
According to our experiments, the running time of this solver is not dependent on the kind of distribution the inputs are sampled from.

@@

DRC is a sequential program generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
The inputs are stored in Pulse Code Modulation (PCM) format, which is an array of 32-bit floating point numbers representing the samples at each sampling time step and the computations are performed using 32-bit precision.
Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform and a DRC-internal implementation of Finite Impulse Response filter generation.
This program accepts user-controllable configurations to decide the width and number of iterations of FFT transforms it uses to perform various passes on the input file, such as windowing, deconvolution, pre- and post-filtering.
The input used in this paper is an audio file of size 768 kilobytes, which is sampled at the rate of 30Khz, 40Khz, ..., 90Khz.

\subsection{Hattrick}
\label{sec:apps:hattrick}
Hattrick is a sequential application simulates the motion of bodies under the effects of gravity using 64-bit precision.
It is designed to help discover extra-solar planets by inferring their existence from Transit Timing Variations, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solve the system's equations of motion.
A given input is described using three parameters: $P$=number of planets, $T$=amount of time to simulate, and $A$=the accuracy target.
In our experiments we considered the following four inputs: $P2T2090A15$, $P2T3090A15$, $P2T4090A15$ and $P3T2090A11$, where an $A15$ and $A11$ denote accuracy targets of $1e-15$ and $1e-11$, respectively.

@@

In this paper we consider a comprehensive resilience strategy that combines various types of error detection and recovery.
Error detection techniques are discussed in Section~\ref{sec:res_tech:err_det}, with Section~\ref{sec:res_tech:err_det:algo} focusing on algorithmic techniques to detect errors in the numeric portions of application state, while Section~\ref{sec:res_tech:err_det:mem} focuses on protection of data structures.
Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart, which offers lower-overhead but higher-latency recovery, as well as replication of key variables, which reduces performance but enables fast recovery.
Finally, Section~\ref{sec:res_tech:eval} presents an experimental evaluation of the performance and accuracy properties of the resilient versions of the major routines used by our target applications, under a wide range of inputs and error injection rates.

@@

\vspace{-5pt}
\subsubsection{Algorithmic Detection}
\label{sec:res_tech:err_det:algo}

@@

Each algorithm used by the above application was enhanced with an algorithm-specific checker that validated its results by exploiting some algorithmic identity.
Each checker computes some aggregate property of the algorithm's results using two different algorithms to produce vectors $v$ and $v'$.
If their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$.
Each error detection causes the routine's execution to be restarted using the rollback mechanism described in Section~\ref{sec:res_tech:cr}.

\vspace{-5pt}
\paragraph{Matrix-matrix multiplication (MMM)}
The operation $A \cdot B$ is checked using a matrix vector multiplication (MVM), via the identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector.
Since we set the check vector $x$ to be all 1s this is equivalent to both sides computing the sums of the columns of matrix $A \cdot B$ using a single MVM on the left and two MVMs on the right.
Most errors in the computation of $A \cdot B$ will affect the sums of its columns and will thus be detected.
This check is efficient because the checker is asymptotically faster than the MMM algorithm, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.

\vspace{-5pt}
\paragraph{Matrix-vector multiplication (MVM)}
The MVM algorithm $Ax=b$ is checked using a similar identity $(x^TA)x = x^T(Ax) = x^Tb$.
We again set $x$ to be the vector of 1's, making the product $x^TA$ equivalent to sum of the columns of matrix $A$.
Although the computation of $x^TA$ involves additions rather than multiplications, the complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM operation.
%While this cost can be reduced significantly by computing $x^TA$ once and reusing it in subsequent multiplications by the same matrix, we did not consider it in this paper because MVM took up little application execution time.
However, for applications such as ADDR where the MVM is applied to same matrix many times with different vectors, the column sums can be reused, amortizing the cost of this computation.
Our experimental evaluation below thus focuses on the overhead of the $x^T(Ax)$ computation, which has $O(n)$ complexity.

\vspace{-5pt}
\paragraph{Symmetric Rank-K update (SYRK)}
The SYRK update is a special case of matrix-matrix multiplication $B = \alpha A \cdot A^T + \beta B$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is incremented into the $n \times n$ matrix $B$.
Its results are checked using the same algorithm as MMM by checking the identity $B x = \alpha A \cdot (A^Tx) + \beta B x$, where $x$ is the vector of all 1's.
This check runs in time $O(n^2)$, as compared to the $O(n^3)$ cost of SYRK.

\vspace{-5pt}
\paragraph{Cholesky Decomposition (CD)}
In this decomposition matrix $A$ is decomposed into $L \cdot L^T$ where $L$ is a lower-triangular matrix with a positive diagonal.
This routine is checked just like MMM by checking the identity $Ax = L \cdot (L^T x)$, which is cheaper than the $O(n^3)$ complexity of the deterministic CD algorithm.
The GSL implements an iterative algorithm that runs faster than $O(n^3)$ time but as shown in the experiments below, it is significantly slower than our checker.

\vspace{-5pt}
\paragraph{Fast Fourier Transform (FFT)}
The FFT algorithm decomposes a function into a sum of sine waves of different frequencies: $f(x) = \sum_{n=0}^{N-1} x_n e^{-i2\pi k n / N}$ for some constant $k$.
We check the results of FFT by using Parseval's theorem, which states that $\sum_{n=0}^{N-1} \left| x[n] \right|^2 = \frac{1}{N} \sum_{k=0}^{N-1} \left| X[k] \right|^2$, where $x$ is the original function and $X$ is its transform. 
Intuitively it means that the energy of the original function is preserved by the transform.
GSL implements optimized radix 2, 3, 4, 5, 6 and 7 subtransform routines as well as a general radix-N FFT routine. The subtransforms are optimized and consume $O(n log(n))$ time and the general one consumes $O(n^2)$ time. Therefore, GSL has a scheduling algorithm that factorizes the length of the input FFT, running the subtransforms whenever possible, using the fallback general radix-N routine only for the remaining factors.
The theorem applies to FFT of any radix and provides an $O(n)$ time check for the $O(n log(n))$ or $O(n^2)$ FFT algorithms.

\vspace{-5pt}
\paragraph{Finite Impulse Response Filter Generation (FIR)}
The FIR filter generation algorithm used in DRC generates a series of samples over the function $sinc(x)=\frac{sin(x)}{x}$
%(\text{if}~x \ne 0)$ and $sinc(x)=1 (\text{if}~x=0)$ 
and modulates it with a Blackman window.
This function has the property that $\int_{-\infty}^{\infty} sinc(x)dx = 1$.
The Blackman window - modulated output of this routine preserves this same property for most of the time, making it possible to check the output by computing the sum and comparing it to 1.
Computing the sum requires $O(n)$ additions, as compared to the $O(n)$ trigonometric function evaluations of the non-iterative FIR generation algorithm.
%As shown below, the overhead of this checker is very small (less than 5\%).
%Since this algorithm has only output and no input and therefore it does not need to have to be reinforced with input backup as other routines do.

\vspace{-5pt}
\paragraph{Runge-Kutta PDE Solver (RK)}
RK is a 4-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
It advances the variable $x$ by steps of size $h$ and uses estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
The implementation of RK in GSL uses adaptive step-size control where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
If the difference between the results obtained with the two step sizes are larger than a threshold $\tau$, the algorithm switches to the smaller step size to maintain accuracy and resumes.
If it is smaller than $\frac{\tau}{2}$, the algorithm switches to a larger step size.
Because soft errors will usually cause significant differences between the two computations, the adaptive step size control algorithm is naturally fault tolerant.
It will react to such errors by rolling back, temporarily increasing its step size and then reverting to the original step size when it observes that this is safe.
As such, we utilize this algorithm to protect the computations within individual RK steps from soft errors.
However, computations in user-provided function derivative functions are not checked since their invariants are notknown.

@@

Since modern systems use 64-bit addressing the range of values that can be assumed by a given pointer is much larger than the amount of memory that can ever be assigned to an application.
This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
Once this address is accessed the resulting violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or Bus Error signal.
This mechanism is cheap but suffers from a long delay between an error and its detection.

An alternative mechanism is to replicate key application variables and to compare the values of the replicas on each access to the variable.
While more expensive, this approach is effective if the number of key variables is small and the cost of rolling back computation is high.

@@

\subsection{Recovery via Checkpoint-Restart and \\Replication}
\label{sec:res_tech:cr}

When errors are detected we employ a recovery method based on checkpoint-restart, where routine state is recorded at key locations and whenever an algorithmic or memory error is detected we roll back to that point in its execution.
Rollbacks are implemented via the \texttt{sigsetjmp}/\texttt{siglongjmp} \cite{sigsetjmp:1997} routines.
\texttt{sigsetjmp} records the application's execution state (registers and stack pointer) at the time of the checkpoint.
On the detection of an error \texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
This method is useful for rolling back to checkpoints recorded in the same function or in a caller function.
When execution returns to the \texttt{sigsetjmp} call the checkpointed portions of the routine's state are recovered and execution repeats.
% The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to negligible performance degradation.

Checkpoint data is protected using two different mechanisms:
Pointers and scalars are replicated such that the checkpoint contains two copies of each.
Arrays of floating point numbers are protected via a block-checksum-based error correcting approach.
By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, it detects the existence of errors and fixes them.
This algorithm takes an array of numbers as input, divides it into blocks of $N \times N$ numbers and for each block it computes and stores the row and column sums, which are used for recovery.

The structure of each routine determines how checkpointing is deployed.
All routines except RK are monolithic code blocks that take a relatively large amount of data as input, operate on it and then return.
In contrast, RK takes in very little input and operates as a long sequence of steps, each of which involves a call to a user-provided derivative routine.
As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers.
The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed.
Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a fixed number of steps between checkpoints.
As our experiments in \ref{sec:res_tech:eval:rk} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.

@@

Having evaluated the resilience properties of the individual application routines and resilience techniques we now consider our three target applications ADDR, Hattrick and DRC.
Section~\ref{sec:eval:perf} reports on the impact of errors and resilience techniques on application performance. Section~\ref{sec:eval:acc} describes the distribution of errors in the results of these applications.
These results are based on a fault injection campaign that includes a range of fault injection rates, input sizes and configurations of resilience mechanisms, with 100 application runs for each combination of parameters.
This corresponds to a total of 66100 for lasso, 43200 for DRC, 102000 for Hattrick error injection experiments.
It was performed on compute nodes with two 2.33 GHz dual-core Xeon E5345 processors with 4GB Ram and RHEL 4 OS.

@@

\begin{figure}[ht!]
\centering
\includegraphics[height=2.75in]{figs/Hattrick_EstdCost.png}
\vspace{-10pt}
\caption{Execution time of Hattrick relative to non-fault tolerant fault-free execution time. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
\vspace{-10pt}
\label{fig:Hattrick_EstdCost}
\end{figure}

@@

\begin{figure}[ht!]
\centering
%\includegraphics[height=3in]{figs/Hattrick_ImperfectRate.png}
\includegraphics[height=2.75in]{figs/Hattrick_ImperfectRate_log.png}
\caption{Probability of not satisfying the user-provided error bound for Hattrick. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
\label{fig:Hattrick_ImperfectRate}
\end{figure}
